Convergent Evolution During Local Adaptation to Patchy Landscapes

### 2015-10-19 patchy-selection-paper-for-PLoS.tex
.

%%%
\subsection{Model of a patchy landscape}
\label{ss:patchyspace}

@@

%%%%%
\subsection[Establishment by Mutation]{Establishment of a locally adaptive allele due to mutational influx}
\label{ss:patchymutation}

@@

%%%%%%%
\subsection[Establishment by Migration]{Establishment of a locally adaptive allele due to migrational influx}
\label{ss:patchymigration}

Now suppose that there are two patches separated by distance $R$
(i.e., the shortest distance between the two is $R$). \revpoint{2}{8}
If the $B$ allele has arisen and become established in the first patch, but has not yet appeared in the second,
we would like to know the time scale on which copies of $B$ move between patches by migration
(supposing that no $B$ allele arises independently by mutation in the meantime).
To determine this, 
we study the fine-scale genealogy of alleles that transit between patches,
obtaining along the way other useful information about the genealogy of $B$ alleles. 
Doing this we arrive at equation \eqref{eqn:migrate} for the rate at which 
an allele established in one patch spreads to a neighboring one by migration.

@@

These asymptotics fit simulations quite well, as shown in Fig~\ref{fig:sim_occupation_freqs}.
To be clear about the assumptions implicit here,
below we provide a derivation of \eqref{eqn:eqfreq} in section \secref{apx:eqfreq},  \revpoint{3}{5}
as well as justification for the asymptotics below in section \secref{apx:asymptotics}.
In one dimension, the equation can be solved to give $q(x)$ 
in terms of Jacobi elliptic functions \citep{NEQwiki}; 
see the Supporting Information.

@@

\begin{figure}[ht!]
  \begin{center}
    \includegraphics{sim-snapshots}
  \end{center}
  \caption{
      \textbf{A segment of the temporal dynamics of the same simulation as Fig~\ref{fig:sim_occupation_freqs}B.}
  Space is again shown across the horizontal axis, with time on the vertical axis; 
  demes are colored darker red the greater their allele frequency at the corresponding time.
  The 50\% and 5\% frequency contour lines are in grey,
  and the genealogy of 50 individuals in a rare long-distance excursion is traced back with black lines.
  }   \label{fig:sim_snapshots}
\end{figure}

@@

Subfamilies typically live for $1/s_m$ generations,
thus spreading across an area of width $\sigma/\sqrt{s_m}$.
The oldest subfamily in a surviving migrant family 
is of order $\log(1/s_m)/s_m$ generations,
so the entire family is spread slightly wider.
\revpoint{2}{2}
% If $\P\{ T_k > k \} = e^{-sk}$
% then 
% \begin{align}
%   \P\{ max_{k \ge n} (T_k - k) \le 0 \} &\simeq \exp(-e^{-sn}/s) \\
%       &= \exp( - \exp( -s(n+\log(s)/s) ) ) \\
% \end{align}
% which is Gumbel with mean $(\log(s)+\gamma)/s$ and variance $\pi^2/(6s^2)$.

\begin{figure}[ht!!]
  \begin{center}
    \includegraphics{branching-concept}
  \end{center}
  \caption{
      \textbf{Cartoon of the decomposition described in the text.}
      \textbf{(a)} 
    Migrants families that get further than $r_0$ from the already-established patch (grey circle)
    are each composed of a ``trunk'' lineage (in red), whose spatial motion we follow,
    embellished by transient branches.
    For the depicted long-lived family, these are further distinguished by those still alive at time $t$ (in black)
    and those that did not survive until $t$ (in grey).
    Fig~\ref{fig:sim_snapshots} has a depiction of such a family, from a simulation. 
    Note that the spatial motions of lineages are independent random walks
    (and hence should be much wigglier than in this cartoon, which confounds space and time).
      \textbf{(b)} 
      Diagram of the area $A'$ used in equation \eqref{eqn:migrate}:
      $R$ is the shortest distance from the old patch to the new,
      and $A'$ is the area of the portion of the new patch that is no more than distance $R+\sigma/\sqrt{2s_m}$
      from the old patch.
} \label{fig:branching_decomp}
\end{figure}

@@

%%%%%%%%%
\subsection{A review, and a test, of the assumptions}
\label{ss:assumptions}

@@

To be more precise, the mutation rate results most obviously apply if $s_m \approx s_p$,
if the patch diameter is large relative to $\sigma/\sqrt{\min(s_m,s_p)}$,
and the patch is not too far from circular.
If $s_m \ll s_p$ then a strip of width $\sigma/\sqrt{s_m}$
should be added to the outside of the patch in computing $A$ for equation \eqref{eqn:mutrate},
and if $s_m \gg s_p$, a strip of width $\sigma/\sqrt{s_p}$ should be subtracted from the patch. \revpoint{2}{6}
As for the migration rate,
we assume that 
each patch is large enough to support a stable population of $B$ alleles
($A^{1/d} > (\sigma/\sqrt{2s_p}) \tan^{-1}(\sqrt{s_m/s_p})$).
The geometry and size of the patch will also affect the approximation of equation \eqref{eqn:gA_1D}.
Next, in using equation \eqref{eqn:eqfreq}, we assume that 
the inter-patch distance is large relative to the characteristic length
($R > \sigma/\sqrt{s_m}$),
and that
local drift is not so strong that the equilibrium frequency is at least approximately attained
(using Wright's effective neighborhood size, $s_m \gg 1/(\rho \sigma^d)$).
The last requirement is necessary because if the $B$ allele fixes in large neighborhoods where it is deleterious,
we cannot approximate its dynamics via a branching process.
We also neglect the time for migration-selection equilibrium to be reached.
As discussed above, we also assume that migration to a new patch takes place over a number of generations;
if there are sufficient rare, long-distance migration events
that would move between patches in a single hop,
this would require a different analysis.

@@

\begin{figure}[ht!]
  \begin{center}
      % made in analyze-sims.R
      \includegraphics{times-predicted-observed}
  \end{center}
  \caption{
      \textbf{Observed versus predicted time until adaptation.}
      \textbf{(left)} 
      Time until adaptation by mutation in an isolated patch for different combinations
      of selection coefficient on patches ($s_m$), 
      population density ($\rho$),
      mutation rate ($\mu$),
      and distance between patches ($R$),
      as observed in simulations (vertical axis),
      and as predicted from equation~\eqref{eqn:mutrate} (horizontal axis).
      In all simulations, $s_p=0.01$.
      \textbf{(right)} 
      Time until adaptation by migration from an already occupied patch to a new patch,
      as observed in simulations (vertical axis),
      and as predicted from equation~\eqref{eqn:migrate}, with $C=5$ (horizontal axis).
      For each,
      parameter combinations and numbers of simulations are in supplemental tables \ref{stab:mutation_params} and \ref{stab:migration_params};
      shown are the median (points) and interquartile range (lines)
      of the time until the patch had at least 100 adapted individuals,
      for each unique parameter combination.
      Only parameter combinations to which the theory here is applicable were used
      (precisely, if
      $R > \sigma/\sqrt{s_m}$, 
      patch width larger than $\sigma/\sqrt{s_m} \atan(\sqrt{s_m/s_p})$, 
      and $s_m > 2 \sigma \rho$);
      the remaining combinations are shown in 
      S\ref{sfig:sim_migration_times} and S\ref{sfig:sim_mutation_times} Fig.
  }   \label{fig:sim_times}
\end{figure}

%%%%%%%%%
\subsection[Probability of Parallel Adaptation]{The probability of parallel adaptation between patches.} 
\label{ss:probparallel}

@@

We can go beyond these rough calculations to find
the probability of parallel adaptation if we are willing to take our approximations at face value. 
Doing so, the time until the first of the two patches is colonized by the $B$
allele will be approximately exponentially distributed with mean $1/(2
\mutrate)$.
Following this, the time until the second patch is subsequently colonized 
(via either migration or new mutation) 
will be exponentially distributed with mean $1/(\mutrate+\migrate)$.
Both these rates scale linearly with population density ($\rho$) 
and the probability of establishment of a rare allele ($p_e\approx 2 s_b/\xi^2$, for $p_e>s_m$),
so that the overall time to adaptation is robustly predicted to increase with offspring variance ($\xi^2$)
and decrease with population density and beneficial selection coefficient.
Furthermore, the probability that the second adaptation is a new mutation,
i.e., the probability of parallel adaptation, 
with $\gamma = \min(1,s_m/p_e)$,
is 
\begin{equation} \label{eqn:parallel_prob}
    \frac{\mutrate}{\mutrate+\migrate(R)} = \frac{ w \mu / (2s_m) }{ w \mu / (2s_m) + C \gamma \left(R \sqrt{2 s_m} /\sigma \right)^{-(d-1)/2}\exp\left(- \sqrt{2 s_m} R / \sigma \right) },  
\end{equation}
so that the probability of parallel mutation should increase
approximately logistically with the distance between the patches, on the spatial scale $\sigma/\sqrt{2 s_m}$. \revpoint{2}{13}

@@

%%%%%%%%%
\subsection[Haplotypes Shared Between Patches]{Length of the hitchhiking haplotype}
\label{ss:haplotype_length}

@@

To keep the expressions simple, in the remainder of this section we only compute quantities for one dimension.
By Bayes' rule, 
\begin{equation} \label{eqn:haplen_cdf}
    % \E[e^{-\ell \tau}|\tau<\infty]  =
    \P\{ L > \ell \}
    = 
    \exp\left\{{-\frac{R}{\sigma}\left(\sqrt{2(\ell+s_m)} - \sqrt{2s_m}\right)}\right\} .
\end{equation}
This can be differentiated to find that the expected transit time is
\begin{equation} 
  \E[\tau \given \tau<\infty ] = (R\sqrt{2s_m}/\sigma)\times 1/(2s_m) \label{eqn:mean_tau}
\end{equation}
and that $\var[\tau \given \tau<\infty ] = (R\sqrt{2s_m}/\sigma) \times 1/(2s_m)^2$.
(A saddle point approximation provides an alternative route to these expressions.) \revpoint{1}{4}

@@

% In Appendix~\ref{apx:limiting_dist_of_HH_haplotype}, 
% we also give some expressions for the limiting distribution of the length of shared haplotype.

% 
% \section{Limiting distribution of the transit time and hitchhiking haplotype}
% \label{apx:limiting_dist_of_HH_haplotype}
% 
% In the above we already assumed that the inter-patch distance $R$ is large relative to the migration-selection cline width $\sigma/\sqrt{s_m}$;
% also taking the large-$R$ limit, obtains a more intuitive description of the process
% via a central limit theorem for $\tau$.
% Define $\eta_\tau$ to be the deviation of $\tau$ from its mean, normalized by its standard deviation,
% so that $\tau = R/\sigma(2s_m)^{-1/2} + \eta_\tau \sqrt{R/\sigma} (2s_m)^{-3/4}$.
% Plugging this into equation~\eqref{eqn:haplen_cdf},
% after some algebra (and using that $\sqrt{R+\epsilon} = \sqrt{R} + \frac{\epsilon}{2\sqrt{R}} - \frac{\epsilon^2}{2\sqrt{R}^3} + O(\epsilon^3)$) we obtain that
% \begin{align}
%   \E[e^{-\ell \eta}] &= \exp\left\{ \frac{\ell^2}{2} + O\left(\frac{1}{R}\right) \right\}.
% \end{align}
% Since the Laplace transform uniquely determines a distribution \citep{durrett1996probability},
% this implies that $\eta_\tau$ converges in distribution to a standard Gaussian random variable as $R \to \infty$.
% In other words, we now know that the distribution of $\tau$ about its mean is approximately Gaussian.
% Since the standard deviation of $\tau$ is of lower order than its mean,
% a similar calculation shows that for large $R$, the haplotype halflength $L$ is approximately exponentially distributed, 
% with mean $\sqrt{2s_m} \sigma/R$.

% \begin{align}
%   \E[e^{-\ell \eta}] &= \E\left[ \exp\left\{-\ell\left(\tau - \frac{R}{\sigma\sqrt{2s_m}}\right)\frac{(2s_m)^{3/4}\sqrt{\sigma}}{\sqrt{R}} \right\} \right] \\
%   &= \E\left[ \exp\left\{ - \left(\frac{\ell (2s_m)^{3/4} \sqrt{\sigma}}{\sqrt{R}}\right) \tau \right\} \right] \exp\left( (2s_m)^{1/4} \sqrt{\frac{R}{\sigma}} \right) \\
% \end{align}

%%%%%%%%%%
\subsection{Applications} 
\label{ss:applications}
% FOR REFERENCE:
% On Wed, May 28, 2014 at 10:20 AM, Michael Nachman <mnachman@berkeley.edu> wrote:
% > Graham:
% >
% > The density is probably in the range of 10-40 animals per hectare (10,000
% > m^s).  We built enclosures of 0.25 hectare and found that they could support
% > about 10 animals with food provided ad libitum, so I think this is an upper
% > density (i.e., 40/hectare).  As for dispersal distance, I'm less sure.
% > Studies on closely related species (genus Perognathus) in similar habitats
% > suggest distances of 100's of feet.  Attached are some old papers that tried
% > to estimate dispersal distances in Perognathus.
% > best,
% > Michael
%
% 
% On Mon, Aug 31, 2015 at 1:42 PM, Michael Nachman <mnachman@berkeley.edu> wrote:
% > Peter:
% >
% > I think this is better, but I would even be a little more cautious, e.g.:
% > "SOME of the largest patches of
% > dark rock range from 10km to 100km wide and 50-400km from each other, and
% > melanic populations of C. intermedius have been found on many of these
% > (but not all) [Benson, Dice & Blossom]. However, patches of all sizes
% > occur across all scales in a heterogeneous manner, so it is not clear
% > to what degree these should be treated as isolated from each other."
% >
% >
% > Note that I changed the numbers a little.  For example, Armendaris and
% > Carrizozo are less than 100 km from each other, and both are big.

Coat color in the rock pocket mouse
\emph{Chaetodipus intermedius} is a classic example of local
adaptation 
\citep{benson1933concealing,DiceBlossom1937}.
These mice live on rocky outcrops throughout the southwestern United States and northern Mexico, 
and generally have a light pelage similar to the predominant rock color.
However, in some regions these mice live on darker substrates (e.g., old lava flows),
and in these patches have much darker pigmentation, 
presumably to avoid visual predators.
Some of the largest patches of dark rock range from 10km to 100km wide
and lie 50--400km from each other,
and dark-colored populations of \emph{C.\ intermedius} have been found on many of these.
(However, patches of all sizes occur across all scales in a heterogeneous manner.)
\revpoint{2}{7}
\citet{nachman2003genetic} demonstrated that on one of these flows (Pinacate), 
much of the change in coloration is due to an allele at the MC1R locus.
This dark allele differs from the light allele by four amino acid changes, 
and has a dominant or partially dominant effect depending on the measure of coat color.
The Pinacate allele is not present in a number of other populations with dark pelage,
suggesting these populations have adapted in parallel \citep{nachman2003genetic,hoekstra2003different}.
However, \citet{hoekstra2004local} reasoned that, elsewhere in the range, 
multiple close dark outcrops may share a dark phenotype whose genetic basis has been spread by migration
despite intervening light habitat.

@@

The base probability of reproduction in each generation in simulations for type $b$ alleles was $r=0.3$;
this was then multiplied by $1+s$ to get the probability of reproduction for type $B$,
where the value of $s$ is either $s_m$ or $s_p$ depending on the individual's location.
This determines the values of $s_m$ and $s_p$ reported in the figures, 
and do not depend on the basic rate of reproduction. \revpoint{2}{5}
However, to obtain values for $s_p$ and $s_m$ when comparing theory to simulation,
we computed the rate of intrinsic growth, 
i.e., the $s$ so that the numbers of $B$ alleles when rare would change by $e^{st}$ after $t$ generations
in the absence of migration.
(The resulting values are close to the first notion of $s$,
but give better agreement with theory,
which uses the second definition.)

@@

\subsection[Probability of Establishment]{Numerical calculation of the probability of establishment}
\label{apx:establishment_sims}

@@

\begin{figure}[ht!]
    \begin{center}
        % made in patchy-selection-plots.R
        \includegraphics{prob-establishment}
    \end{center}
    \caption{
        \textbf{The probability of establishment of a new mutant allele as a function of distance from the edge of the region where it is beneficial,
        in an abrupt transition (left) and a gradual transition (right).}
    The allele is deleterious to the left and beneficial to the right (with selection coefficients $\mp 0.05$ respectively);
    and the selection coefficient interpolates linearly between these in the central hatched region.
    The number of offspring is Poisson.
    The underlying landscape has demes every 15km and a migration rate of 0.5 between neighboring demes;
    the probability was found by numerically solving the equation for the probability of establishment of a multi-type branching process.
    Note that the approximation used in the text is here $p_e \approx 2 s_p / \xi^2 = 0.1$,
    agreeing with the value in the center of the patch.
    \label{fig:prob_estab_calcs}
    }
\end{figure}

%%%%%%%%%%%%
\subsection[Equilibrium Frequency]{The equilibrium frequency}
\label{apx:eqfreq}

@@

In one dimension, we are done (and discuss exact solutions in the Supporting Information);
in higher dimensions, we are more interested in the mean frequency at a given distance $r$ from a patch.
If we take a radially symmetric patch centered at the origin (so $s$ only depends on $r$), 
and let $\xi(t,r)$ denote the mean occupation frequency at distance $r$,
then the polar form of the Laplacian in $d$ dimensions gives us that \eqref{eqn:derived_fkpp} is
\begin{align}
  \label{eqn:radial_fkpp}
  \partial_t \xi(t,r) = \mu \left( \frac{\sigma^2}{2} \partial_{r}^2 \xi(t,r) + \sigma^2\frac{d-1}{2r} \partial_r \xi(t,r) + s(r) \xi(t,r) (1-\xi(t,r)) \right) .
\end{align}

%%%%%%%%%%%
\subsection[Asymptotics]{Asymptotic solution for the equilibrium frequency} 
\label{apx:asymptotics}

@@

%%%%%%%%% %%%%%%%%%%%%%%
\subsection[Hitting and Occupation]{Hitting and occupation}
\label{ss:hitting_occupation}

@@

%%%%%%% %%%%%%%%%%%%
\subsection[The Integral $q(S)$]{The integral $q(S)$}
\label{apx:qS}

@@

\section*{Acknowledgements}
We would like to heartily thank Joachim Hermisson, Richard Neher, and an anonymous reviewer
for a number of insightful suggestions,
and catching several errors present in earlier drafts. \revpoint{3}{4}
We would also like to acknowledge Michael Nachman, Steve Evans, and Yaniv Brandvain for useful discussions, 
as well as Gideon Bradburd, and Simon Aeschbacher and the rest of the Coop lab
for helpful comments and discussion of the manuscript.

>>>>

%%%
\subsection*{Model of a patchy landscape}
\label{ss:patchyspace}

@@

%%%%%
% \subsection[Establishment by Mutation]{Establishment of a locally adaptive allele due to mutational influx}
\subsection*{Establishment of a locally adaptive allele due to mutational influx}
\label{ss:patchymutation}

@@

%%%%%%%
% \subsection[Establishment by Migration]{Establishment of a locally adaptive allele due to migrational influx}
\subsection*{Establishment of a locally adaptive allele due to migrational influx}
\label{ss:patchymigration}

Now suppose that there are two patches separated by distance $R$
(i.e., the shortest distance between the two is $R$).
If the $B$ allele has arisen and become established in the first patch, but has not yet appeared in the second,
we would like to know the time scale on which copies of $B$ move between patches by migration
(supposing that no $B$ allele arises independently by mutation in the meantime).
To determine this, 
we study the fine-scale genealogy of alleles that transit between patches,
obtaining along the way other useful information about the genealogy of $B$ alleles. 
Doing this we arrive at equation \eqref{eqn:migrate} for the rate at which 
an allele established in one patch spreads to a neighboring one by migration.

@@

These asymptotics fit simulations quite well, as shown in Fig~\ref{fig:sim_occupation_freqs}.
To be clear about the assumptions implicit here,
below we provide a derivation of \eqref{eqn:eqfreq} in section \secref{apx:eqfreq},
as well as justification for the asymptotics below in section \secref{apx:asymptotics}.
In one dimension, the equation can be solved to give $q(x)$ 
in terms of Jacobi elliptic functions \citep{NEQwiki}; 
see the Supporting Information.

@@

\begin{figure}[ht!]
  \begin{center}
    % \includegraphics{sim-snapshots}
  \end{center}
  \caption{
      \textbf{A segment of the temporal dynamics of the same simulation as Fig~\ref{fig:sim_occupation_freqs}B.}
  Space is again shown across the horizontal axis, with time on the vertical axis; 
  demes are colored darker red the greater their allele frequency at the corresponding time.
  The 50\% and 5\% frequency contour lines are in grey,
  and the genealogy of 50 individuals in a rare long-distance excursion is traced back with black lines.
  }   \label{fig:sim_snapshots}
\end{figure}

@@

Subfamilies typically live for $1/s_m$ generations,
thus spreading across an area of width $\sigma/\sqrt{s_m}$.
The oldest subfamily in a surviving migrant family 
is of order $\log(1/s_m)/s_m$ generations,
so the entire family is spread slightly wider.
% If $\P\{ T_k > k \} = e^{-sk}$
% then 
% \begin{align}
%   \P\{ max_{k \ge n} (T_k - k) \le 0 \} &\simeq \exp(-e^{-sn}/s) \\
%       &= \exp( - \exp( -s(n+\log(s)/s) ) ) \\
% \end{align}
% which is Gumbel with mean $(\log(s)+\gamma)/s$ and variance $\pi^2/(6s^2)$.

\begin{figure}[ht!!]
  \begin{center}
    % \includegraphics{branching-concept}
  \end{center}
  \caption{
      \textbf{Cartoon of the decomposition described in the text.}
      \textbf{(a)} 
    Migrants families that get further than $r_0$ from the already-established patch (grey circle)
    are each composed of a ``trunk'' lineage (in red), whose spatial motion we follow,
    embellished by transient branches.
    For the depicted long-lived family, these are further distinguished by those still alive at time $t$ (in black)
    and those that did not survive until $t$ (in grey).
    Fig~\ref{fig:sim_snapshots} has a depiction of such a family, from a simulation. 
    Note that the spatial motions of lineages are independent random walks
    (and hence should be much wigglier than in this cartoon, which confounds space and time).
      \textbf{(b)} 
      Diagram of the area $A'$ used in equation \eqref{eqn:migrate}:
      $R$ is the shortest distance from the old patch to the new,
      and $A'$ is the area of the portion of the new patch that is no more than distance $R+\sigma/\sqrt{2s_m}$
      from the old patch.
} \label{fig:branching_decomp}
\end{figure}

@@

%%%%%%%%%
\subsection*{A review, and a test, of the assumptions}
\label{ss:assumptions}

@@

To be more precise, the mutation rate results most obviously apply if $s_m \approx s_p$,
if the patch diameter is large relative to $\sigma/\sqrt{\min(s_m,s_p)}$,
and the patch is not too far from circular.
If $s_m \ll s_p$ then a strip of width $\sigma/\sqrt{s_m}$
should be added to the outside of the patch in computing $A$ for equation \eqref{eqn:mutrate},
and if $s_m \gg s_p$, a strip of width $\sigma/\sqrt{s_p}$ should be subtracted from the patch.
As for the migration rate,
we assume that 
each patch is large enough to support a stable population of $B$ alleles
($A^{1/d} > (\sigma/\sqrt{2s_p}) \tan^{-1}(\sqrt{s_m/s_p})$).
The geometry and size of the patch will also affect the approximation of equation \eqref{eqn:gA_1D}.
Next, in using equation \eqref{eqn:eqfreq}, we assume that 
the inter-patch distance is large relative to the characteristic length
($R > \sigma/\sqrt{s_m}$),
and that
local drift is not so strong that the equilibrium frequency is at least approximately attained
(using Wright's effective neighborhood size, $s_m \gg 1/(\rho \sigma^d)$).
The last requirement is necessary because if the $B$ allele fixes in large neighborhoods where it is deleterious,
we cannot approximate its dynamics via a branching process.
We also neglect the time for migration-selection equilibrium to be reached.
As discussed above, we also assume that migration to a new patch takes place over a number of generations;
if there are sufficient rare, long-distance migration events
that would move between patches in a single hop,
this would require a different analysis.

@@

\begin{figure}[ht!]
  \begin{center}
      % made in analyze-sims.R
      % \includegraphics{times-predicted-observed}
  \end{center}
  \caption{
      \textbf{Observed versus predicted time until adaptation.}
      \textbf{(left)} 
      Time until adaptation by mutation in an isolated patch for different combinations
      of selection coefficient on patches ($s_m$), 
      population density ($\rho$),
      mutation rate ($\mu$),
      and distance between patches ($R$),
      as observed in simulations (vertical axis),
      and as predicted from equation~\eqref{eqn:mutrate} (horizontal axis).
      In all simulations, $s_p=0.01$.
      \textbf{(right)} 
      Time until adaptation by migration from an already occupied patch to a new patch,
      as observed in simulations (vertical axis),
      and as predicted from equation~\eqref{eqn:migrate}, with $C=5$ (horizontal axis).
      For each,
      parameter combinations and numbers of simulations are in supplemental tables \ref{stab:mutation_params} and \ref{stab:migration_params};
      shown are the median (points) and interquartile range (lines)
      of the time until the patch had at least 100 adapted individuals,
      for each unique parameter combination.
      Only parameter combinations to which the theory here is applicable were used
      (precisely, if
      $R > \sigma/\sqrt{s_m}$, 
      patch width larger than $\sigma/\sqrt{s_m} \atan(\sqrt{s_m/s_p})$, 
      and $s_m > 2 \sigma \rho$);
      the remaining combinations are shown in 
      S\ref{sfig:sim_migration_times} and S\ref{sfig:sim_mutation_times} Fig.
  }   \label{fig:sim_times}
\end{figure}

%%%%%%%%%
% \subsection[Probability of Parallel Adaptation]{The probability of parallel adaptation between patches.} 
\subsection*{The probability of parallel adaptation between patches.} 
\label{ss:probparallel}

@@

We can go beyond these rough calculations to find
the probability of parallel adaptation if we are willing to take our approximations at face value. 
Doing so, the time until the first of the two patches is colonized by the $B$
allele will be approximately exponentially distributed with mean $1/(2
\mutrate)$.
Following this, the time until the second patch is subsequently colonized 
(via either migration or new mutation) 
will be exponentially distributed with mean $1/(\mutrate+\migrate)$.
Both these rates scale linearly with population density ($\rho$) 
and the probability of establishment of a rare allele ($p_e\approx 2 s_b/\xi^2$, for $p_e>s_m$),
so that the overall time to adaptation is robustly predicted to increase with offspring variance ($\xi^2$)
and decrease with population density and beneficial selection coefficient.
Furthermore, the probability that the second adaptation is a new mutation,
i.e., the probability of parallel adaptation, 
with $\gamma = \min(1,s_m/p_e)$,
is 
\begin{equation} \label{eqn:parallel_prob}
    \frac{\mutrate}{\mutrate+\migrate(R)} = \frac{ w \mu / (2s_m) }{ w \mu / (2s_m) + C \gamma \left(R \sqrt{2 s_m} /\sigma \right)^{-(d-1)/2}\exp\left(- \sqrt{2 s_m} R / \sigma \right) },  
\end{equation}
so that the probability of parallel mutation should increase
approximately logistically with the distance between the patches, on the spatial scale $\sigma/\sqrt{2 s_m}$.

@@

%%%%%%%%%
% \subsection[Haplotypes Shared Between Patches]{Length of the hitchhiking haplotype}
\subsection*{Length of the hitchhiking haplotype}
\label{ss:haplotype_length}

@@

To keep the expressions simple, in the remainder of this section we only compute quantities for one dimension.
By Bayes' rule, 
\begin{equation} \label{eqn:haplen_cdf}
    % \E[e^{-\ell \tau}|\tau<\infty]  =
    \P\{ L > \ell \}
    = 
    \exp\left\{{-\frac{R}{\sigma}\left(\sqrt{2(\ell+s_m)} - \sqrt{2s_m}\right)}\right\} .
\end{equation}
This can be differentiated to find that the expected transit time is
\begin{equation} 
  \E[\tau \given \tau<\infty ] = (R\sqrt{2s_m}/\sigma)\times 1/(2s_m) \label{eqn:mean_tau}
\end{equation}
and that $\var[\tau \given \tau<\infty ] = (R\sqrt{2s_m}/\sigma) \times 1/(2s_m)^2$.
(A saddle point approximation provides an alternative route to these expressions.)

@@

%%%%%%%%%%
\subsection*{Applications} 
\label{ss:applications}
% FOR REFERENCE:
% On Wed, May 28, 2014 at 10:20 AM, Michael Nachman <mnachman@berkeley.edu> wrote:
% > Graham:
% >
% > The density is probably in the range of 10-40 animals per hectare (10,000
% > m^s).  We built enclosures of 0.25 hectare and found that they could support
% > about 10 animals with food provided ad libitum, so I think this is an upper
% > density (i.e., 40/hectare).  As for dispersal distance, I'm less sure.
% > Studies on closely related species (genus Perognathus) in similar habitats
% > suggest distances of 100's of feet.  Attached are some old papers that tried
% > to estimate dispersal distances in Perognathus.
% > best,
% > Michael
%
% 
% On Mon, Aug 31, 2015 at 1:42 PM, Michael Nachman <mnachman@berkeley.edu> wrote:
% > Peter:
% >
% > I think this is better, but I would even be a little more cautious, e.g.:
% > "SOME of the largest patches of
% > dark rock range from 10km to 100km wide and 50-400km from each other, and
% > melanic populations of C. intermedius have been found on many of these
% > (but not all) [Benson, Dice & Blossom]. However, patches of all sizes
% > occur across all scales in a heterogeneous manner, so it is not clear
% > to what degree these should be treated as isolated from each other."
% >
% >
% > Note that I changed the numbers a little.  For example, Armendaris and
% > Carrizozo are less than 100 km from each other, and both are big.

Coat color in the rock pocket mouse
\emph{Chaetodipus intermedius} is a classic example of local
adaptation 
\citep{benson1933concealing,DiceBlossom1937}.
These mice live on rocky outcrops throughout the southwestern United States and northern Mexico, 
and generally have a light pelage similar to the predominant rock color.
However, in some regions these mice live on darker substrates (e.g., old lava flows),
and in these patches have much darker pigmentation, 
presumably to avoid visual predators.
Some of the largest patches of dark rock range from 10km to 100km wide
and lie 50--400km from each other,
and dark-colored populations of \emph{C.\ intermedius} have been found on many of these.
(However, patches of all sizes occur across all scales in a heterogeneous manner.)
\citet{nachman2003genetic} demonstrated that on one of these flows (Pinacate), 
much of the change in coloration is due to an allele at the MC1R locus.
This dark allele differs from the light allele by four amino acid changes, 
and has a dominant or partially dominant effect depending on the measure of coat color.
The Pinacate allele is not present in a number of other populations with dark pelage,
suggesting these populations have adapted in parallel \citep{nachman2003genetic,hoekstra2003different}.
However, \citet{hoekstra2004local} reasoned that, elsewhere in the range, 
multiple close dark outcrops may share a dark phenotype whose genetic basis has been spread by migration
despite intervening light habitat.

@@

The base probability of reproduction in each generation in simulations for type $b$ alleles was $r=0.3$;
this was then multiplied by $1+s$ to get the probability of reproduction for type $B$,
where the value of $s$ is either $s_m$ or $s_p$ depending on the individual's location.
This determines the values of $s_m$ and $s_p$ reported in the figures, 
and do not depend on the basic rate of reproduction.
However, to obtain values for $s_p$ and $s_m$ when comparing theory to simulation,
we computed the rate of intrinsic growth, 
i.e., the $s$ so that the numbers of $B$ alleles when rare would change by $e^{st}$ after $t$ generations
in the absence of migration.
(The resulting values are close to the first notion of $s$,
but give better agreement with theory,
which uses the second definition.)

@@

% \subsection[Probability of Establishment]{Numerical calculation of the probability of establishment}
\subsection*{Numerical calculation of the probability of establishment}
\label{apx:establishment_sims}

@@

\begin{figure}[ht!]
    \begin{center}
        % made in patchy-selection-plots.R
        % \includegraphics{prob-establishment}
    \end{center}
    \caption{
        \textbf{The probability of establishment of a new mutant allele as a function of distance from the edge of the region where it is beneficial,
        in an abrupt transition (left) and a gradual transition (right).}
    The allele is deleterious to the left and beneficial to the right (with selection coefficients $\mp 0.05$ respectively);
    and the selection coefficient interpolates linearly between these in the central hatched region.
    The number of offspring is Poisson.
    The underlying landscape has demes every 15km and a migration rate of 0.5 between neighboring demes;
    the probability was found by numerically solving the equation for the probability of establishment of a multi-type branching process.
    Note that the approximation used in the text is here $p_e \approx 2 s_p / \xi^2 = 0.1$,
    agreeing with the value in the center of the patch.
    \label{fig:prob_estab_calcs}
    }
\end{figure}

%%%%%%%%%%%%
% \subsection[Equilibrium Frequency]{The equilibrium frequency}
\subsection*{The equilibrium frequency}
\label{apx:eqfreq}

@@

In one dimension, we are done (and discuss exact solutions in S1 Text);
in higher dimensions, we are more interested in the mean frequency at a given distance $r$ from a patch.
If we take a radially symmetric patch centered at the origin (so $s$ only depends on $r$), 
and let $\xi(t,r)$ denote the mean occupation frequency at distance $r$,
then the polar form of the Laplacian in $d$ dimensions gives us that \eqref{eqn:derived_fkpp} is
\begin{align}
  \label{eqn:radial_fkpp}
  \partial_t \xi(t,r) = \mu \left( \frac{\sigma^2}{2} \partial_{r}^2 \xi(t,r) + \sigma^2\frac{d-1}{2r} \partial_r \xi(t,r) + s(r) \xi(t,r) (1-\xi(t,r)) \right) .
\end{align}

%%%%%%%%%%%
% \subsection[Asymptotics]{Asymptotic solution for the equilibrium frequency} 
\subsection*{Asymptotic solution for the equilibrium frequency} 
\label{apx:asymptotics}

@@

%%%%%%%%% %%%%%%%%%%%%%%
% \subsection[Hitting and Occupation]{Hitting and occupation}
\subsection*{Hitting and occupation}
\label{ss:hitting_occupation}

@@

%%%%%%% %%%%%%%%%%%%
% \subsection[The Integral $q(S)$]{The integral $q(S)$}
\subsection*{The integral $q(S)$}
\label{apx:qS}

@@

\subsection*{Text S1}

\section*{Acknowledgements}
We would like to heartily thank Joachim Hermisson, Richard Neher, and an anonymous reviewer
for a number of insightful suggestions,
and catching several errors present in earlier drafts. 
We would also like to acknowledge Michael Nachman, Steve Evans, and Yaniv Brandvain for useful discussions, 
as well as Gideon Bradburd, and Simon Aeschbacher and the rest of the Coop lab
for helpful comments and discussion of the manuscript.
